Military mechanization as a complex adaptive system

### 2018-07-29 paper2/paper2.tex
foo

BEEF THIS UP AND YOU'RE DONE

The model results in much higher
system-level mechanization than occurred in reality. This higher
mechanization is not evenly distributed across agents. Rather, a small number of
agents with many enemies and that all border each other become outliers and
hence drive the global average far upward. The explanation for both phenomena
lies in a positive feedback loop. Enemies and neighbors adjust their
mechanization scores in response to each other; hence, a geographic cluster of
enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
mechanization.

Because this model is a faithful representation of the factors that \citet{sechser2010army} claim are
most important for mechanization, that model output fails to match reality
indicates that their claimed model is incomplete. There must be some ``brake''
on mechanization; put differently, some countervailing factor(s) must mitigate the
effect of the positive feedback loop identified above. This model cannot
conclusively identify what that ``brake'' is. The most plausible candidates are
factors examined by \citet{sechser2010army} and deemed statistically
insignificant such as regime type and economy, and factors not examined by
\citet{sechser2010army} at all such as culture.

This paper is a novel contribution to the literature. Although agent-based
models have found traction in the study of grievance, insurgency, and
other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
they quite rarely are used to model international relations and state-centric security
phenomena despite some early work on this front
\citep[see][]{,axelrod1997complexity}. 
This paper's agentization of an orthodox security
studies model thus is novel in method. Its findings also are important. They
show that they show a solely ``external'', security
environment-focused view of mechanization does not stand up to agentization.
Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
loops---all hallmarks of a complex system---drive results that significantly
diverge from theoretical expectations of the orthodox view.

@@

>>>>

The model results in much higher system-level mechanization than occurred in reality. This higher
mechanization is not evenly distributed across agents. Rather, a small number of
agents with many enemies and that all border each other become outliers and
hence drive the global average far upward. The explanation for both phenomena
lies in a positive feedback loop. Enemies and neighbors adjust their
mechanization scores in response to each other; hence, a geographic cluster of
enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their mechanization
and drive system-level mechanization upwards. Modeled sources of anti-mechanization (e.g.,
defeat in counter-insurgency) are insufficient to hold mechanization to historical levels.

Because this model is a faithful representation of the factors that the orthodox
literature claim are most important for mechanization, that model output fails to match reality
indicates that the orthodox view of mechanization is incomplete. It must not be
near-solely driven by external security factors. There must be some additional ``brake''
on mechanization. Put differently, some countervailing factor(s) must mitigate the
effect of the positive feedback loop identified above. This model cannot conclusively 
identify what that ``brake'' is, but preliminary extensions to the model provide
some insight, as described in the \textbf{Future Research} section.

The model's behavior recreates IR phenomena from the bottom
up. The aforementioned positive feedback loop is well-studied in IR as the
``security dilemma'' \citep{jervis1978cooperation}. Note that there is no model
logic explicitly dictating this behavior. Rather, the model endows agents with 
only simple rules---e.g., \textit{if my neighbor increases mechanization, 
then increase my own}---that nevertheless cause the emergence of this behavior.
That it does so increases confidence in the correctness of the model since
it in this specific area corresponds to theoretically-expected behavior despite
not being programmed to do so.

More broadly, model behavior supports the argument that mechanization is
a complex system. There is no definitive checklist for what constitutes
complexity, but this model's behavior meets at least four core criteria. One,
as previously described, agent cognition and interactions cause behavior 
to emerge in a bottom-up fashion. Two, because the model's macro-behavior derives from
micro-interactions, one cannot causally separate the part from the whole. For
example, suppose that State A increases its mechanization, and therefore neighboring
State B also does so. One cannot say that State A \textit{caused} State B to
raise its mechanization since State A's original increase itself derives from
interaction with other neighbors, enemies, historical experiences, and the like. 
Three, a positive feedback loop emerges. Four, system behavior is in some ways 
resilient to variation in input parameter values. As shown in Table 6's parameter sweep, very
different input values lead to about the same behavior: e.g., the same states
are always relatively the most mechanized. That the model's behavior matches so
many attributes of complex systems strengthens the belief that real-world
mechanization is a complex system (put differently, that these findings are not
just artifacts of a particular model).

@@

This paper's findings raise an obvious and important question: what is the
missing ``brake'' that, if included in the model, 
would hold mechanization to historical levels? Fully answering this question is
beyond the scope of this paper.
